---
title: "Project1"
author: "Cassie, Caleb, & Mike"
date: "Monday, February 02, 2015"
output: html_document
---
**In this project, we will document and explain the steps taken in order to take data from a database and make interesting graphs.**

First, we need to load the libraries "RCurl", used for processing HTTP requests, and "reshape2", used to restructure data.

Next, we call the customers table, found using the query:

```
select * from customers
```

Once we have assigned the customers table data to the ```dfc``` dataframe, we can view the first few rows by calling the function:

```
print(head(dfc))
```

These steps are shown below.

```{r}

#source("../01 Data/dataframes.R", echo = TRUE)

source("../01 Data/dfc.R", echo = TRUE)

```

**Now, let's do something fun with the data.** First, we need to load the "ggplot2" library, which is a plotting system for R.

As we can see, there are many columns to this database: ```CUSTOMER_ID```, ```CUSTOMER_FIRST_NAME```, ```CUSTOMER_LAST_NAME```, ...

Let's make a plot of this data that displays the area code versus city for each customer, coloring each point by state. It's not a particularly useful graph, but it conveys the ability of R to readily display certain data in a graphical format.

Note that the ```ggplot()``` function accepts the dataframe as its first argument, followed by the  ```aes()``` function as its second argument to define aesthetics of the graph area.

The call to make a scatterplot is:

```
geom_point(mapping = NULL, data = NULL, stat = "identity", position = "identity", na.rm = FALSE, ...)
```

This function is added to the ```ggplot()``` function call to define the graph as a scatterplot.

The R code that creates this graph is shown below.

```{r}

source("../02 Visualizations/dfcplot.R", echo = TRUE)

```

Next, we'll make a dataframe from the order_details table, using the query:

```
select * from order_details
```

```{r}

source("../01 Data/dfd.R", echo = TRUE)

```

This just has the columns ```ORDER_ID```, ```ITEM_ID```, and ```ORDER_QTY``` - not particularly exciting. Also, all but two of the ```ORDER_QTY``` rows are equal to 1. So, let's add some qualifiers to our search query to spice this data up:

```
select ITEM_ID, sum(ORDER_QTY) Total_Quantity from order_details group by ITEM_ID order by ITEM_ID
```

Here, we have ignored the ```ORDER_ID``` column and instead grouped the data by the ```ITEM_ID```, with the corresponding value equal to the total number of orders for each item - ```sum(ORDER_QTY)```.

```{r}

source("../01 Data/dfdSum.R", echo = TRUE)

```

Now that we have the data grouped and ordered by item, we can make a bar graph showing the total quantity purchased of each item. The call to make a bar graph is:

```
geom_bar(mapping = NULL, data = NULL, stat = "bin", position = "stack", ...)
```

```{r} 

source("../02 Visualizations/dfdSumplot.R", echo = TRUE)

```

Moving on to the orders table, we set the search query to:

```
select * from orders
```

We can see that the columns are ```ORDER_ID```, ```CUSTOMER_ID```, ```ORDER_DATE```, and ```SHIPPED_DATE```.

```{r}

source("../01 Data/dfo.R", echo = TRUE)

```

From this data, we can make a scatterplot showing the ```ORDER_ID``` versus the ```CUSTOMER_ID```, with the color of each point representative of the ```CUSTOMER_ID```. The ```as.factor()``` function separates the colors into distinct groups, as shown in the legend to the right of the graph. Without this function, the colors would follow a gradient along the domain of ```CUSTOMER_ID```.

```{r}

source("../02 Visualizations/dfoplot.R", echo = TRUE)

```

Similarly, the search query for the items table is:

```
select * from items
```

This table has the columns ```ITEM_ID```, ```TITLE```, ```ARTIST```, and ```UNIT_PRICE```.

```{r}

source("../01 Data/dfi.R", echo = TRUE)

```

Now, we'll make a scatterplot of the data: ```UNIT_PRICE``` versus ```ARTIST```.

```{r}

source("../02 Visualizations/dfiplot.R", echo = TRUE)

```

Let's find a way to aggregate the data from all four tables and make a graph that reveals something about the entire data set.

We'll use the search query:

```
select * from orders o join customers c on o.customer_id = c.customer_id join order_details d on o.order_id = d.order_id join items i on d.item_id = i.item_id
```

In this query, we can see that there is the typical "select all" call in the form of ```select * from orders```. This is followed by three ```join``` clauses, which define how to connect the tables together. For instance, ```o.customer_id = c.customer_id``` joins the tables by connecting the customer ID from each table and adding the other corresponding values to each row. The same technique joins the order ID from the order_details table, and the item ID from the items table.

```{r}

source("../01 Data/dfm.R", echo = TRUE)

```

Finally, lets plot the ```SHIPPED_DATE``` versus the ```ORDER_DATE``` for the orders, separating the scatterplots by ```CUSTOMER_STATE```. To separate the graphical area into separate plots for each state, use the ```facet_wrap()``` function with the customer state as the argument.

The ```as.Date()``` function allows the user to specify the formatting of each date as it appears on each axis.

We'll create this graph three times, using ```UNIT_PRICE```, ```TITLE```, and ```ARTIST``` as the color variables, respectively.

```{r}

source("../02 Visualizations/dfmplot.R", echo = TRUE)

```

```{r}

#library("ggplot2")
#library("reshape2")

#source("../02 Visualizations/Graphs.R", echo = TRUE)

```