---
title: "Project1"
author: "Cassie &  Caleb"
date: "Monday, February 02, 2015"
output: html_document
---
**In this project, we will document and explain the steps taken in order to take data from a database and make interesting graphs.**

First, we need to load the libraries "RCurl", used for processing HTTP requests, and "reshape2", used to restructure data.

Next, we call the "customers" table, found using the query:

```
select * from customers
```

Once we have assigned the "customers" table data to the ```dfc``` dataframe, we can view the first few rows by calling the function:

```
print(head(dfc))
```

These steps are shown below.

```{r}

#source("../01 Data/dataframes.R", echo = TRUE)

source("../01 Data/dfc.R", echo = TRUE)

```

**Now, let's do something fun with the data.** First, we need to load the "ggplot2" library, which is a plotting system for R.

As we can see, there are many columns to this database: ```CUSTOMER_ID```, ```CUSTOMER_FIRST_NAME```, ```CUSTOMER_LAST_NAME```, ...

Let's make a plot of this data that displays the area code versus city for each customer, coloring each point by state. It's not a particularly useful graph, but it conveys the ability of R to readily display certain data in a graphical format.

The R code that creates this graph is shown below.

```{r}

source("../02 Visualizations/dfcplot.R", echo = TRUE)

```

Next, we'll make a dataframe from the order_details table, using the query:

```
select * from order_details
```

```{r}

source("../01 Data/dfd.R", echo = TRUE)

```

This just has the columns ```ORDER_ID```, ```ITEM_ID```, and ```ORDER_QTY``` - not particularly exciting. Also, all but two of the ```ORDER_QTY``` rows are equal to 1. So, let's add some qualifiers to our search query to spice this data up:

```
select ITEM_ID, sum(ORDER_QTY) Total_Quantity from order_details group by ITEM_ID order by ITEM_ID
```

Here, we have ignored the ```ORDER_ID``` column and instead grouped the data by the ```ITEM_ID```, with the corresponding value equal to the total number of orders for each item - ```sum(ORDER_QTY)```.

```{r}

source("../01 Data/dfdSum.R", echo = TRUE)

```

```{r} 

source("../02 Visualizations/dfdSumplot.R", echo = TRUE)

```

```{r}

source("../01 Data/dfo.R", echo = TRUE)

```

```{r}

source("../02 Visualizations/dfoplot.R", echo = TRUE)

```

```{r}

source("../01 Data/dfi.R", echo = TRUE)

```

```{r}

source("../02 Visualizations/dfiplot.R", echo = TRUE)

```

```{r}

source("../01 Data/dfm.R", echo = TRUE)

```

```{r}

source("../02 Visualizations/dfmplot.R", echo = TRUE)

```

```{r}

#library("ggplot2")
#library("reshape2")

#source("../02 Visualizations/Graphs.R", echo = TRUE)

```